---
output:
  html_document: default
  pdf_document: default
params: 
    set_title: "full" # full transmission 
    file: 'prelim_results/full.with.covid_simset_2024-07-15_C.12580.Rdata'
    file.date: '2024-07-15'
    notes: 'Notes:' 
title: "`r paste(tools::toTitleCase(params$set_title), 'Calibration')`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r source and load files, include=FALSE}
source('applications/EHE/ehe_specification.R')
load(params$file)
```


Calibration **`r simset$calibration.code`** run on **`r params$file.date`**  

**Notes:** `r params$notes`

### Summary Plots

The following plots are included below:  
1. Population (total and by age/race)  
2. New Diagnoses (total, by risk/race and age/sex)  
3. Diagnosed Prevalence (total, by risk/race and age/sex)  
4. Awareness and Test Positivity (totals) 
5. Testing (total and by age, sex, race, and risk)  
6. AIDS diagnoses (total and by age, sex, race, and risk)  
7. Suppression (total and by age, sex, race, and risk)  
8. HIV Mortality, Total Mortality, and AIDS Deaths (total)  
9. PrEP uptake (total, age, sex) and PrEP indications (total, age, sex)  
10.Proportion using heroin/cocaine (total and by age)  
11. Immigration and emigration (total, age, race)  


```{r set up plots, echo=FALSE}
location.style.manager = create.style.manager(color.data.by = "location.type") 
source.style.manager = create.style.manager(color.data.by = "source") 
stratum.style.manager = create.style.manager(color.data.by = "stratum") 
```


### 1. Population
Total and by age/race

```{r population plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
  simplot(simset,
        outcomes = c("population"), 
        style.manager = source.style.manager, 
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
  simplot(simset,
        facet.by = "age", split.by = "race", 
        outcomes = c("population"), 
        dimension.values = list(year = 2000:2030)) )))
```



### 2. New Diagnoses
Total, by risk/race, and by age/sex

```{r new diagnoses plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("new"), 
        style.manager = source.style.manager, 
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "risk", split.by = "race", 
        outcomes = c("new"), 
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", split.by = "sex", 
        outcomes = c("new"), 
        dimension.values = list(year = 2000:2030)) )))
```



### 3. Diagnosed Prevalence
Total, by risk/race, and by age/sex

```{r prevalence plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("diagnosed.prevalence"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "risk", split.by = "race", 
        outcomes = c("diagnosed.prevalence"), 
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", split.by = "sex", 
        outcomes = c("diagnosed.prevalence"), 
        dimension.values = list(year = 2000:2030)) )))

```



### 4. Awareness and Test Positivity 
Totals  

```{r awareness and positivity plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("awareness"), # totals only 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("cdc.hiv.test.positivity"), # totals only 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))
```



### 5. Testing
Total and by age, sex, race, and risk

```{r testing plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("testing"),
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", 
        outcomes = c("testing"),
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "sex", 
        outcomes = c("testing"),
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "race", 
        outcomes = c("testing"),
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "risk", 
        outcomes = c("testing"),
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))
```



### 6. AIDS Diagnoses
Total and by age, sex, race, and risk

```{r aids diagnoses plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("aids.diagnoses"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 1980:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", 
        outcomes = c("aids.diagnoses"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 1980:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "sex", 
        outcomes = c("aids.diagnoses"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 1980:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "race", 
        outcomes = c("aids.diagnoses"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 1980:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "risk", 
        outcomes = c("aids.diagnoses"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 1980:2030)) )))
```



### 7. Suppression
Total and by age, sex, race, and risk

```{r suppression plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("suppression"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", 
        outcomes = c("suppression"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "sex", 
        outcomes = c("suppression"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "race", 
        outcomes = c("suppression"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "risk",
        outcomes = c("suppression"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))
```



### 8. HIV Mortality, Total Mortality, and AIDS Deaths
Totals

```{r mortality plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        #facet.by = "sex", # sex; 1-way 
        outcomes = c("hiv.mortality"),
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("total.mortality"), # totals only 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

# must ask for years 1981-2001 
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("aids.deaths"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 1981:2001)) )))

```



### 9. PrEP
Uptake by total, age and sex
Indications by total, age, sex

```{r prep plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("prep.uptake"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", 
        outcomes = c("prep.uptake"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "sex", 
        outcomes = c("prep.uptake"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("prep.indications"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", # age, sex; 1-way 
        outcomes = c("prep.indications"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "sex", # age, sex; 1-way 
        outcomes = c("prep.indications"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

```



### 10. IDU
Proportion using heroin/cocaine; total and by age

```{r IDU plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("proportion.using.heroin",
                     "proportion.using.cocaine"), 
        style.manager = source.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", # age; 1-way 
        outcomes = c("proportion.using.heroin",
                     "proportion.using.cocaine"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

```



### 11. Migration
Immigration and emigration; total and by age and race

```{r migration plots, echo=FALSE}
suppressWarnings(invisible(capture.output(
simplot(simset,
        outcomes = c("immigration","emigration"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "age", # age, race; 1-way 
        outcomes = c("immigration","emigration"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

suppressWarnings(invisible(capture.output(
simplot(simset,
        facet.by = "race", # age, race; 1-way 
        outcomes = c("immigration","emigration"), 
        style.manager = location.style.manager,
        dimension.values = list(year = 2000:2030)) )))

```
